Supplementary Information

### 2020-08-26 doc/manuscript/MEE_manuscript_SI.Rmd
added site figures to SI, SI formatting

---
title: "Supplementary Information"
author:
output: 
  pdf_document:
    fig_caption: yes

@@

\listoftables
\listoffigures

@@

@@

```{r eval = TRUE, echo=FALSE, warning=FALSE}
library(knitr)
library(kableExtra)
S1 <- read.csv("tables_figures/site_details.csv", stringsAsFactors = FALSE, check.names = FALSE)
kable(S1, format = "latex", booktabs = TRUE,  caption = "Table S1. Site details", escape = F)%>%
  kable_styling(latex_options = c("scale_down", "hold_position"), protect_latex = T) 
```

\newpage
```{r eval = TRUE, echo=FALSE, warning=FALSE}
library(knitr)
library(kableExtra)
S2 <- read.csv("tables_figures/species.csv", stringsAsFactors = FALSE, check.names = FALSE)
kable(S2, format = "latex", booktabs = TRUE, caption = "Table S2. List of species analyzed", escape = F)%>%
  kable_styling(latex_options = c("scale_down", "hold_position")) 
```

** Table S3- allometric equations for bark thickness **

@@

\newpage 
![**Figure S1 | (Comparison of traditional approaches with ours).**  (THIS FIGURE IS JUST A MOCK-UP TO SHOW VALENTINE WHAT I HAVE IN MIND.) ](tables_figures/mock_comparison_traditional_method.png)

![**Figure S2 | Example comparison of climwin output when the optimal climate variable and time window were identical across growth metrics.** Example is temperature variable group at Little Tesuque. Here, *climwin* identified potential evapotranspiration ($PET$) as the strongest climate variable across all three metrics of growth ($\Delta r$, BAI, $\Delta AGB$) and regardless of whether all cores were included in the analysis, or only those for which DBH could be reconstructed ($\Delta r$-trees with $DBH$, BAI, $\Delta AGB$).)](/Users/kteixeira/Dropbox (Smithsonian)/GitHub/EcoClimLab/ForestGEO-climate-sensitivity/results/climwin_plots_combined/NewMexico_pet.png)

![**Figure S3 | Example comparison of climwin output when the optimal climate variable differed across growth metrics, but the selected time window was similar.** Example is precipitation variable group at Little Tesuque. Here, *climwin* identified precipitation (PRE) as the strongest climate variable for $\Delta r$ and $BAI$, but precipitation day frequency (WET) as the strongest climate variable for $\Delta AGB$.](/Users/kteixeira/Dropbox (Smithsonian)/GitHub/EcoClimLab/ForestGEO-climate-sensitivity/results/climwin_plots_combined/NewMexico_wet_pre.png)

![**Figure S4 | Example comparison of climwin output when both the optimal climate variable and selected time window differed across growth metrics.** Example is precipitation variable group at Harvard Forest. Here, *climwin* identified precipitation frequency (WET) as the strongest climate variable for $\Delta r$, but precipitation amount (PRE) as the strongest climate variable for $BAI$ and $\Delta AGB$. The optimal time window (circled) also differed across growth metrics.](/Users/kteixeira/Dropbox (Smithsonian)/GitHub/EcoClimLab/ForestGEO-climate-sensitivity/results/climwin_plots_combined/HarvardForest_pre_wet.png)

>>>>

---
title: "Supplementary Information"
author:
output: 
  pdf_document:
    fig_caption: yes
    toc: true # table of content true
    toc_depth: 3

@@

@@

\newpage
## Tables S1. Site Details
```{r eval = TRUE, echo=FALSE, warning=FALSE}
library(knitr)
library(kableExtra)
S1 <- read.csv("tables_figures/site_details.csv", stringsAsFactors = FALSE, check.names = FALSE)
kable(S1, format = "latex", booktabs = TRUE,   escape = F)%>%
  kable_styling(latex_options = c("scale_down", "hold_position"), protect_latex = T) 
```

\newpage
## Tables S2. List of species analyzed
```{r eval = TRUE, echo=FALSE, warning=FALSE}
library(knitr)
library(kableExtra)
S2 <- read.csv("tables_figures/species.csv", stringsAsFactors = FALSE, check.names = FALSE)
kable(S2, format = "latex", booktabs = TRUE,  escape = F)%>%
  kable_styling(latex_options = c("scale_down", "hold_position")) 
```

@@

## Table S3. allometric equations for bark thickness

@@

\newpage 
## Figure S1. Comparison of our approach with traditional methods of identifying climate signals
![**Figure S1 | (Comparison of traditional approaches with ours).**  (THIS FIGURE IS JUST A MOCK-UP TO SHOW VALENTINE WHAT I HAVE IN MIND.) ](tables_figures/mock_comparison_traditional_method.png){ width=70% }

\newpage
## Figure S2. Comparison of climwin output across growth metrics for the temperature variable group at Little Tesuque (New Mexico, USA)
![**Figure S2 | Comparison of climwin output across growth metrics for the temperature variable group at Little Tesuque (New Mexico, USA).** Here, *climwin* identified potential evapotranspiration ($PET$) as the strongest climate variable across all three metrics of growth ($\Delta r$, BAI, $\Delta AGB$) and regardless of whether all cores were included in the analysis, or only those for which DBH could be reconstructed ($\Delta r$-trees with $DBH$, BAI, $\Delta AGB$).)](/Users/kteixeira/Dropbox (Smithsonian)/GitHub/EcoClimLab/ForestGEO-climate-sensitivity/results/climwin_plots_combined/NewMexico_pet.png)

\newpage
## Figure S3. Comparison of climwin output across growth metrics for the precipitation variable group at Little Tesuque (New Mexico, USA)
![**Figure S3 | Comparison of climwin output across growth metrics for the precipitation variable group at Little Tesuque (New Mexico, USA).** Here, *climwin* identified precipitation (PRE) as the strongest climate variable for $\Delta r$ and $BAI$, but precipitation day frequency (WET) as the strongest climate variable for $\Delta AGB$.](/Users/kteixeira/Dropbox (Smithsonian)/GitHub/EcoClimLab/ForestGEO-climate-sensitivity/results/climwin_plots_combined/NewMexico_wet_pre.png)

\newpage
## Figure S4. Comparison of climwin output across growth metrics for the precipitation variable group at Harvard Forest (Massachusetts, USA)
![**Figure S4 | Comparison of climwin output across growth metrics for the precipitation variable group at Harvard Forest (Massachusetts, USA).** Here, *climwin* identified precipitation frequency (WET) as the strongest climate variable for $\Delta r$, but precipitation amount (PRE) as the strongest climate variable for $BAI$ and $\Delta AGB$. The optimal time window (circled) also differed across growth metrics.](/Users/kteixeira/Dropbox (Smithsonian)/GitHub/EcoClimLab/ForestGEO-climate-sensitivity/results/climwin_plots_combined/HarvardForest_pre_wet.png)

\newpage
## Figure S5. Best GLS models for Barro Colorado Island (Panama)

![**Figure S5 | Best GLS models for Barro Colorado Island (Panama) for all three growth metrics: $\Delta r$, $BAI$, and $\Delta AGB$.** Precipitation and temperature group variables are as selected by *climwin*. For each species, relationships are plotted if included in top model, with dashed lines indicating terms that do not significantly improve the model (relative to a model without). Vertical grey lines indicate the long-term mean for the climate variable, shading indicates 1 SD.](/Users/kteixeira/Dropbox (Smithsonian)/GitHub/EcoClimLab/ForestGEO-climate-sensitivity/results/composite_plots/BCI.png){ width=78% }

\newpage
## Figure S6. Best GLS models for Huai Kha Khaeng (Thailand)

![**Figure S6 | Best GLS models for Huai Kha Khaeng (Thailand) for all three growth metrics: $\Delta r$, $BAI$, and $\Delta AGB$.** Precipitation and temperature group variables are as selected by *climwin*. For each species, relationships are plotted if included in top model, with dashed lines indicating terms that do not significantly improve the model (relative to a model without). Vertical grey lines indicate the long-term mean for the climate variable, shading indicates 1 SD.](/Users/kteixeira/Dropbox (Smithsonian)/GitHub/EcoClimLab/ForestGEO-climate-sensitivity/results/composite_plots/HKK.png){ width=78% }

\newpage
## Figure S7. Best GLS models for Little Tesuque (New Mexico, USA)

![**Figure S7 | Best GLS models for Little Tesuque  (New Mexico, USA) for all three growth metrics: $\Delta r$, $BAI$, and $\Delta AGB$.** Precipitation and temperature group variables are as selected by *climwin*. For each species, relationships are plotted if included in top model, with dashed lines indicating terms that do not significantly improve the model (relative to a model without). Vertical grey lines indicate the long-term mean for the climate variable, shading indicates 1 SD.](/Users/kteixeira/Dropbox (Smithsonian)/GitHub/EcoClimLab/ForestGEO-climate-sensitivity/results/composite_plots/NewMexico.png){ width=78% }

\newpage
## Figure S8. Best GLS models for Cedar Breaks (Utah, USA)

[**Figure S8 | Best GLS models for Cedar Breaks  (Utah, USA) for all three growth metrics: $\Delta r$, $BAI$, and $\Delta AGB$.** Precipitation and temperature group variables are as selected by *climwin*. For each species, relationships are plotted if included in top model, with dashed lines indicating terms that do not significantly improve the model (relative to a model without). Vertical grey lines indicate the long-term mean for the climate variable, shading indicates 1 SD.]

\newpage
## Figure S9. Best GLS models for SCBI (Virginia, USA)

![**Figure S9 | Best GLS models for SCBI (Virginia, USA) for all three growth metrics: $\Delta r$, $BAI$, and $\Delta AGB$.** Precipitation and temperature group variables are as selected by *climwin*. For each species, relationships are plotted if included in top model, with dashed lines indicating terms that do not significantly improve the model (relative to a model without). Vertical grey lines indicate the long-term mean for the climate variable, shading indicates 1 SD.](/Users/kteixeira/Dropbox (Smithsonian)/GitHub/EcoClimLab/ForestGEO-climate-sensitivity/results/composite_plots/SCBI.png){ width=78% }

\newpage
## Figure S10. Best GLS models for Lilly Dickey Woods (Indiana, USA)

![**Figure S10 | Best GLS models for Lilly Dickey Woods (Indiana, USA) for all three growth metrics: $\Delta r$, $BAI$, and $\Delta AGB$.** Precipitation and temperature group variables are as selected by *climwin*. For each species, relationships are plotted if included in top model, with dashed lines indicating terms that do not significantly improve the model (relative to a model without). Vertical grey lines indicate the long-term mean for the climate variable, shading indicates 1 SD.](/Users/kteixeira/Dropbox (Smithsonian)/GitHub/EcoClimLab/ForestGEO-climate-sensitivity/results/composite_plots/LillyDickey.png){ width=78% }

\newpage
## Figure S11. Best GLS models for Harvard Forest (Massachusetts, USA)

![**Figure S11 | Best GLS models for Harvard Forest (Massachusetts, USA) for all three growth metrics: $\Delta r$, $BAI$, and $\Delta AGB$.** Precipitation and temperature group variables are as selected by *climwin*. For each species, relationships are plotted if included in top model, with dashed lines indicating terms that do not significantly improve the model (relative to a model without). Vertical grey lines indicate the long-term mean for the climate variable, shading indicates 1 SD.](/Users/kteixeira/Dropbox (Smithsonian)/GitHub/EcoClimLab/ForestGEO-climate-sensitivity/results/composite_plots/HarvardForest.png){ width=78% }

\newpage
## Figure S12. Best GLS models for Niobrara/ Hansley (Nebraska, USA)

\newpage
## Figure S13. Best GLS models for Zofin (Czech Republic)

![**Figure S13 | Best GLS models for Zofin (Czech Republic) for all three growth metrics: $\Delta r$, $BAI$, and $\Delta AGB$.** Precipitation and temperature group variables are as selected by *climwin*. For each species, relationships are plotted if included in top model, with dashed lines indicating terms that do not significantly improve the model (relative to a model without). Vertical grey lines indicate the long-term mean for the climate variable, shading indicates 1 SD.](/Users/kteixeira/Dropbox (Smithsonian)/GitHub/EcoClimLab/ForestGEO-climate-sensitivity/results/composite_plots/Zofin.png){ width=78% }

\newpage
## Figure S14. Best GLS models for Scotty Creek (NW Territories, Canada)

![**Figure S14 | Best GLS models for Scotty Creek (NW Territories, Canada) for all three growth metrics: $\Delta r$, $BAI$, and $\Delta AGB$.** Precipitation and temperature group variables are as selected by *climwin*. For each species, relationships are plotted if included in top model, with dashed lines indicating terms that do not significantly improve the model (relative to a model without). Vertical grey lines indicate the long-term mean for the climate variable, shading indicates 1 SD.](/Users/kteixeira/Dropbox (Smithsonian)/GitHub/EcoClimLab/ForestGEO-climate-sensitivity/results/composite_plots/ScottyCreek.png){ width=78% }
